---
title: Introduction to R Package Development 2019
author: Stephanie Hicks
date: '2019-05-20'
slug: introduction-to-r-package-development-2019
categories:
  - blog
tags:
  - blog
header:
  caption: ''
  image: ''
  preview: yes
output: 
    html_document:
        theme: cosmo 
        toc: true
        toc_float: true
        highlight: tango
        number_sections: false 
---

# Motivation

In this tutorial, you are going to create your 
own package using the `usethis` workflow R package. 
For our purposes, we are using the code provided on the 
[`usethis`](https://usethis.r-lib.org) website and the 
[blog from Emil Hvitfeldt](https://www.hvitfeldt.me/blog/usethis-workflow-for-package-development/). 

This package is great because it automates repetitive tasks that 
arise during project setup and development, both for 
R packages and non-package projects.

## Getting started with the `usethis` package 

### Installation

Install the released version of `usethis` from CRAN:

```{r, eval=FALSE}
install.packages("usethis")
```

Or install the development version from GitHub with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("r-lib/usethis")
```

You may also need these packages: 

```{r, eval=FALSE}
install.packages("testthat")
install.packages("spelling")
```


# Creating your R package using `usethis`

Most `use_*()` functions operate on the active project:
literally, a directory on your computer. If you have just
used `usethis` to create a new package or project, that
will be the active project. Otherwise, `usethis` verifies
that current working directory is or is below a valid 
project directory and that becomes the active project.

A few `usethis` functions have no strong connections to 
projects and will expect you to provide a path.

Below is a quick look at how `usethis` can help to set up a 
package. But remember, many `usethis` functions are also 
applicable to analytical projects that are not packages.

Let's load the library 

```{r}
library(usethis)
```

Next we create a path to a temporary folder and call the 
package `mypkg`.

```{r}
path <- file.path(tempdir(), "mypkg")
path
```

Then, we use the `create_package()` function 
to create the skeleton of this package. 

## Create the package 

```{r}
create_package(path)
```

If you are inside of RStudio, this will (hopefully)
open up a new RStudio session. Also, `usethis` is
quite chatty, explaining what it’s doing 
and assigning you tasks. For example, 

✔ indicates something usethis has done for you. 

● indicates that you’ll need to do some work yourself.

**For this tutorial, we are not in an interactive session**. 
So we use the `proj_activate(path)` function with the 
`path` argument to change the working directory to the 
temporary path that we've created. 

**NOTE**: You only need to run `proj_activate(path)` 
if you are not in an interactive session. 

```{r}
proj_activate(path)
```

## Setting up git/github

We won't do it here, but you can use the functions 
below to help set link the pakckage to GitHub. If you 
want to explore this, check out 
[Jenny Bryan tutorial on getting setup](https://github.com/r-lib/usethis/blob/master/vignettes/articles/usethis-setup.Rmd). 

```{r, eval=FALSE}
use_git()
use_github()
```

## One time modifications

Now that we are up and running there is a bunch of 
things we should do before we start writing code. 
Firstly we will go over all the actions that only 
have to be done once and get those out of the way.


### Setting up your license

There are several types of licenses you can pick from, 
for example 

```{r, eval=FALSE}
use_mit_license()
use_gpl3_license()
use_apl2_license()
use_cc0_license()
```

Let's select the MIT license 

```{r}
use_mit_license("Margaret Taub")
```

The choice of license is beyond the scope of the tutorial, 
but we [refer readers here](http://r-pkgs.had.co.nz/description.html#license) 
to learn more about types of licenses. 


### Add README.md files 

Now we add the readme files, this is done using
the `use_readme_md()` function. This will create a 
readme.md file that you can edit and knit 
as you normally would.

```{r}
use_readme_md()
```

### Adding continuous integration (optional)

Next we will setup some continuous integration. 
This is completely optional, but it's recommended. 
A nice intro to this topic is from 
[Julia Silge](https://juliasilge.com/blog/beginners-guide-to-travis/). 

For a quick summary, we will extract a section from
her blog post: 

> "If you have written an R package, you probably have 
gotten into the habit of running R CMD check on a 
(hopefully) regular basis to check for common problems. 
It automatically detects lots of problems that can arise 
that you might not otherwise notice. Also you probably 
have put your package on GitHub, for version control and 
also to be able to share your code with other people and
track issues and whatnot. The idea behind continuous 
integration is that CI will automatically run R CMD check 
(along with your tests, etc.) every time you push a commit
to GitHub. You don’t have to remember to do this; CI 
automatically checks the code after every commit. Travis-CI
is a continuous integration service that builds and tests 
projects that are hosted on GitHub; if your R package is 
open source (i.e., not private), then using Travis-CI is free." 

To set up CI, use the following `usethis` functions: 

```{r, eval=FALSE}
use_travis()
use_appveyor()
use_coverage(type = c("codecov"))
```

Again, this is outside the scope of this tutorial, but we 
leave it here for you to explore. 

### Add unit testing

You will most likely also want to include 
unit testing, this can be achieved using the 
[`testthat`](https://cran.r-project.org/package=testthat)
package. To do this, use the `use_testthat()` function. 

```{r}
use_testthat()
use_test("my-test")
```

Additional information about testing can be found [`here`](http://r-pkgs.had.co.nz/tests.html).

### Add spell check
Next we will add spell checking to our workflow, this is done with
```{r}
# install.packages("spelling")
use_spell_check()
```

Make sure that the spelling package is installed before running.

### Create data folder and objects
If you are going to include data in your package, you 
would want to include a data-raw folder where the data
is created/formatted.

```{r}
use_data_raw()

x <- 1
y <- 2
use_data(x, y)
```

### Add NEWS file

Lastly if you plan on doing a little larger project a 
NEWS file is very handy to keep track on what is
happening in your package.

```{r}
use_news_md()
```

### Add packages to DESCRIPTION file
You can also add packages to the DESCRIPTION file 

For example, here we are adding the `MASS` package 
to the Suggests description field. 

```{r}
use_package("MASS", "Suggests")
use_package("dplyr")
```

### Set up documentation files 

We can set up the documention files using Roxygen

```{r}
use_roxygen_md()
```

(You may need to install the `roxygen` package first.)

## Multiple time modifications

Now that we have setup all the basics, 
the general development can begin.

You typical workflow will be repeating the following
steps in the order that suits your flow

* Write some code
* Restart R Session Cmd+Shift+F10 (Ctrl+Shift+F10 for Windows)
* Build and Reload Cmd+Shift+B (Ctrl+Shift+B for Windows)
* Test Package Cmd+Shift+T (Ctrl+Shift+T for Windows)
* Check Package Cmd+Shift+E (Ctrl+Shift+E for Windows)
* Document Package Cmd+Shift+D (Ctrl+Shift+D for Windows)

### Create `.R` files and unit tests for functions

```{r}
use_r("function_name")
```

This function is very important and you will using it a lot, 
not only will it create the files you save your functions 
in, but it will also open the files if they are already 
created, this makes navigating your R files much easier. 
Once you have created your function it is time to add 
some tests! This is done using the `use_test()` function,
and it works much the same way as the `use_r()`.

```{r}
use_test("function_name")
```

### Create a vignette 

An vignette provides a nice piece of documentation once you have added a bunch of capabilities to your package.

```{r}
use_vignette("my_cool_vignette")
```

# Summary

You should be off and running now! 

